Have you forgotten?' Conversion experiences in lists

### 2016-10-19 lists_conversion_experience.rmd
writing abstract

---
title: "'Have you forgotten?' Conversion experiences in list worlds"
author: "Adrian Mackenzie"
date: "`r Sys.Date()`"
output:
tufte::tufte_handout: default
citation_package: biber
latex_engine: xelatex
bibliography: ~/ref_bibs/uni.bib
link-citations: yes
---

# 'Have you forgotten?' Conversion experiences in list worlds

@@

Shopping lists such as Figure \ref{fig:shopping_list} are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people. I regard the list of five suggestions show to Tesco's online customers at the checkout as symptomatic of a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.

In this respect, the presentation was not really surprising or exceptional. The ordinariness and prosaic character of grocery shopping stands out as the most significant feature.  The presentation itself is a working example for an argument about our experience of predictive ordering. In this paper, I re-present the presentation using a mixture of theoretical and practical resources. The theory is largely recent media theory and its accounts of prediction in media. The practical resources are that of coding and predictive models.

The problem of incomplete lists reminds me of the game 'I went shopping and I bought ... ' that adults sometimes play with children.  The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.  The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items. The longer the game continues, the harder it is to remember since the shopping list grows longer. At the same time, each player repeats the list, and this  helps memorisation. One trick in playing this game is to choose items that are quite unexpected or increasingly random. Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children. The children have better memories for items they know, but adults know more unexpected items. On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children. Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated. Something similar applies to recommendation systems. The system can remember very many things. It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.

A recommendation system should suggest items that you have not forgotten but may have never bought. If you do buy recommended items, then the recommender system has been effective in shunting the situation differently. It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high-value. There is little value in recommending a whole lot of small and cheap items such as rubber bands or toothpicks. So it might need to choose amongst different recommendations and weight high value items differently to low value items. In either case, 'conversion'  is the valued event.

We could regard lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA]. In this paper, I focus on lists and their propensities in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].  List practices such as recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.

**How do we approach lists as objects of research?** While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhere refractory to discourse analysis and textual analysis methods of humanities and social science research. For on the one hand, lists are highly associative. They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings. For on the other hand, as operational entities, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...  shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]

@@

It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer. It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured. The lift or uplift refers to a change in events associated with the same groups of people. This effectively is an experiment in inhabiting two different worlds.  This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^12] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift. Predictions, then, are not bound by fixed parameters of the model.

[^12]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models. Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000. See for instance [@Almquist_2001]  Response modelling, propensity modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.  Interestingly, these

@@

![propensity_matrix](figure/propensity_conversions.png)

>>>>

---
title: "'Have you forgotten?' Conversion experiences in lists"
author: "Adrian Mackenzie"
date: "`r Sys.Date()`"
output:
tufte::tufte_handout: default
citation_package: biber
latex_engine: xelatex
bibliography: ~/ref_bibs/uni.bib
link-citations: yes
---

# 'Have you forgotten?' Conversion experiences in lists

@@

Shopping lists such as Figure \ref{fig:shopping_list} are quite prosaic -- a dozen lines scrawled on the back of a used envelope stuck on the refrigerator door  -- and sublime -- a million products ranked in some kind of order for tens or hundreds of millions of people. Such lists are highly repetitive, yet mix items of great diversity and display complicated temporal structures, spanning past and future, actual and potential. I regard the list of five suggestions show to Tesco's online customers at the checkout as exemplifying a major contemporary form of ordering, the recommendation list, that we see in many  forms of news, media, entertainment, marketing, advertising.

The presentation I saw was not really surprising or exceptional. It was a presentation by a data scientist to data scientists, describing an implementation of a predictive model as well as some of the problems and potentials of the predictive model.  I see the presentation itself as providing materials for an argument about our experience of predictive ordering. In this paper, I re-present the presentation using a mixture of theoretical and practical resources. The theory is largely recent media theory and its accounts of prediction in media. The practical resources are that of coding and predictive models.[^12]

[^12]: The problem of incomplete lists reminds me of the game 'I went shopping and I bought ... ' that adults sometimes play with children.  The game is about remembering a well-ordered list of items - 'a' is apple, 'b' is for banana, 'c' is for carrot.  The point of the game is to use the well-ordered set of labels provided by the sequence of letters in the alphabet to remember a random and perhaps very diverse list of items. The longer the game continues, the harder it is to remember since the shopping list grows longer. At the same time, each player repeats the list, and this  helps memorisation. One trick in playing this game is to choose items that are quite unexpected or increasingly random. Rather than `ham`, list `halva`; rather than `fish`, add `face cream.` This is especially useful for adults in beating children. The children have better memories for items they know, but adults know more unexpected items. On the other hand, very unexpected items are quite easy to remember and tend to even the score in the favour of the children. Balancing between the improbable and memorable, between a predictable list like 'apples, bananas, carrots' and an unusual list such as 'agave syrup, boccadilla, cilantro' is complicated. Something similar applies to recommendation systems. The system can remember very many things. It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.

A recommendation system should suggest items that you have not forgotten but may have never bought. If you do buy recommended items, then the recommender system has been effective in shunting the situation differently. It has led to a **'conversion.'** From a supermarket perspective, the conversion should be high in price. There is little value in successfully recommending a whole lot of small and cheap items such as rubber bands or toothpicks. So it might need to choose amongst different recommendations and weight high value items differently to low value items. In either case, 'conversion'  is the valued event.

**How do we approach lists as objects of research?** While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhat refractory to discourse  and textual analysis methods. On the one hand, lists are highly associative. They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that multiply semiotic readings. On the other hand, as operational inscriptions, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...  shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]

List practices such as online grocery recommendations figure as sites of intersection between people, infrastructures, things, places and predictions.  We might, therefore,  approach lists from various novel angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@Lury_2012]. In this paper, I focus on lists and list-making in terms of the predictive conditions of contemporary media experience [@Hansen_2015a].

@@

It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer. It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured. The lift or uplift refers to a change in events associated with the same groups of people. This effectively is an experiment in inhabiting two different worlds.  This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^122] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift. Predictions, then, are not bound by fixed parameters of the model.

[^122]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models. Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000. See for instance [@Almquist_2001]  Response modelling, propensity modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.  Interestingly, these

@@

![propensity_matrix](figure/propensity_conversions.jpg)
